---
output: github_document
---

# TL;DR

 - Our group focused on Sale Price as the main variable of interest.

- The distribution of sale prices is right-skewed, with most homes sold between $150,000–$300,000, a few low-value outliers under $50,000, and a handful of very high-value sales above $600,000.

- We explored TotalLivingArea (sf) as a related variable. Its distribution is also right-skewed, with most homes under 2,000 sq ft and a few very large properties above 5,000 sq ft.

- Scatterplots show a positive relationship between living area and sale price: generally, larger homes sell for more.

- Outliers remain: some small homes with unusually high prices (likely due to land value or location) and some very large homes with surprisingly low prices (possibly condition or neighborhood effects).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(classdata)


library(dplyr)
library(ggplot2)

```



```{r}

colnames(ames)
str(ames)
head(ames)
max(ames$`Sale Price`)

```

As a team, we noticed the two categorical/numeric variables, for example, some numeric variables are Sale Price, Year Built, Acres, TotalLivingArea, and various others. The categorial variables, such as AC, Fireplace, and Multi Sale, indicate presence/absence features, while numeric variables capture continuous measures like price and footage. We predict the sales price to range from 0 to multiple millions of dollars.


# Step 2

- Our variable of interest is the sale price.

# Step 3

```{r}
# Summary of Sale Price
summary(ames$'Sale Price')

library(ggplot2)
salePriceGraph <- ggplot(ames, aes(x = `Sale Price`)) + 
  geom_histogram(binwidth = 50000, fill = "skyblue", color = "black")

print(salePriceGraph)

```


The histogram shows that most homes are clustered around or below $500,000. Some oddities are present, certain values are 0 and a few outliers were homes sold for millions but not very many overall.

``` {r}
boxplot(ames$`Sale Price`,
        horizontal = TRUE,
        main = "Boxplot of Sale Prices")
```

The boxplot shows the skewed distribution and as well as the outliers, which confirms what we saw in the histogram above.

# Step 3

``` {r}

plot(ames$`TotalLivingArea (sf)`, ames$`Sale Price`,
     main = "Sale Price vs. Living Area",
     xlab = "Total Living Area (sf)",
     ylab = "Sale Price",
     pch  = 20,
     col  = "darkgreen")

```

# Step 4

  - Nathan Krieger
  
    ``` {r}
    
    ames_clean <- ames %>%
      filter(!is.na(`Sale Price`), `Sale Price` > 0,
             !is.na(`TotalLivingArea (sf)`), `TotalLivingArea (sf)` > 0)
    
   # Range of TotalLivingArea (sf)
  range(ames_clean$`TotalLivingArea (sf)`)
  
  # Range of Sale Price
  range(ames_clean$`Sale Price`)
      
    # Histogram of TotalLivingArea
    ggplot(ames, aes(x = `TotalLivingArea (sf)`)) +
      geom_histogram(binwidth = 250, fill = "lightgreen", color = "black") +
      labs(title = "Distribution of Total Living Area",
           x = "Total Living Area (sq ft)",
           y = "Number of Homes") +
      theme_minimal()
    
    # Scatterplot: Sale Price vs TotalLivingArea
    ggplot(ames, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
      geom_point(alpha = 0.6, color = "steelblue") +
      labs(title = "Sale Price vs Total Living Area",
           x = "Total Living Area (sq ft)",
           y = "Sale Price (USD)") +
      theme_minimal()
    
    ```
    - The var I chose is TotalLivingArea (sf). I cleaned the data by dropping all 0 and NaN values for SQFT and sale price. I found that the overall distribution of SQFT is about what is expected, a right skewed plot. There are some homes with SQFT near 6000, but most are below 2000. Based off of my second plot, there appears to be a loose correlation between SQFT and price. The plot is mostly flat, with some outliers that have very high prices but low SQFT. I assume these are houses/properties with lots of land.
    
    
  - Phillip
    ``` {r}
    library(ggplot2)

    # make a clean subset using base R
    ames_clean <- subset(ames,
                         !is.na(`Sale Price`) & `Sale Price` > 0 &
                         !is.na(`TotalLivingArea (sf)`))
    
    livingAreaPlot <- ggplot(ames_clean,
                             aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
      geom_point(color = "darkgreen", alpha = 0.5, size = 1.5) +
      labs(title = "Sale Price vs. Total Living Area",
           x = "Total Living Area (sf)",
           y = "Sale Price")
    
    print(livingAreaPlot)

    ```
    
    Phillip Giametta: I chose the variable Total Living Area (sF), after dropping the 0/NA prices, we can see a clear positive relationship where larger homes sell for more money. We can see evidence of a few large homes selling for lower than expected, which could be for various reasons (location, condition etc).


